---
title: "Ecology and Economy in the sea: strengths and gaps in trans-dominion communication"
subtitle: "R notebook accompanying the paper"
output:
  html_document:
    theme: sandstone
    highlight: tango
    code_folding: hide
    fig_width: 10
---

```{r include=FALSE}
  library("tidyverse")
  library("kableExtra")
  library("textmineR")
  library("igraph")
  library("alluvial")
  library("ggdendro")
  library("tidygraph")
  library("ggraph")
  library("gridExtra")
  library("dendextend")
  Sys.setlocale("LC_ALL", "C")
```

### Loading and tidying the data
The original file is based on an [isi/wos](https://www.webofknowledge.org/) search on the query "marine AND ecology AND economy" (~1000 hits).
After removing all the entries without abstract (~300 hits), we screened the abstracts for inconsistencies, removing ~200 articles. ~600 articles remain.
```{r loadata}
  dataset <- read.csv(file = "./inputs/dataset per analisi.csv",
                      header = T, sep = ";", stringsAsFactors = F)
```

We select title and abstract from the file, use the ISI identifier for the id column, and join title and abstract into a single row. Later we discover that many abstracts end with copyright notices, so we remove those.
```{r tidydat}
  dataset.clean <- dataset %>%
    select(UT,TI,AB) %>%
    as_tibble %>%
    unite(TI_and_ABS,TI,AB, sep = " ") %>%
    rename(intText=TI_and_ABS,id=UT) %>%
    mutate(intText= gsub("\\([Cc]) [0-9]... .*","",intText))
```

# Using textmineR
We choose the [textimineR](https://cran.r-project.org/web/packages/textmineR) package for topic modelling because its topic naming and testing capabilities.

We create the document term matrix (rows are documents, columns are single words) and keep all the words found more than five times in the corpus (i.e. the document collection).
```{r dtmcr, message=FALSE, results='hide'}
  dtm <- CreateDtm(doc_vec = dataset.clean$intText, # character vector of documents
                   doc_names = dataset.clean$id, # document names
                   ngram_window = c(1, 2), # minimum and maximum n-gram length
                   stopword_vec = c(stopwords::stopwords("en"), # stopwords from tm
                                    stopwords::stopwords(source = "smart")), # this is the default value
                   lower = TRUE, # lowercase - this is the default value
                   remove_punctuation = TRUE, # punctuation - this is the default
                   remove_numbers = TRUE, # numbers - this is the default
                   verbose = TRUE)

  dtm <- dtm[, colSums(dtm > 0) > 5] #Here we select words appearing more than five times
```

## Basic corpus statistic
Now that we have the document/term matrix, we can calculate term frequency (_how often each term appears in the dtm_), document frequency (_in how many documents the term appears_), and inverse document frequency, for both terms and bigrams.
```{r basicor, echo=FALSE, message=FALSE}
  tf_mat <- TermDocFreq(dtm = dtm)
  tf_bigrams <- tf_mat[stringr::str_detect(tf_mat$term, "_"), ]

  tfh  <- head(tf_bigrams[order(tf_bigrams$term_freq, decreasing = TRUE), ], 10)
  tfbh <- head(tf_mat[order(tf_mat$term_freq, decreasing = TRUE), ], 10)
  kable(tfh, escape =F, align="c") %>%
   kable_styling(bootstrap_options = c("striped", "hover"), full_width=F, position = "float_left", font_size=12)
  kable(tfbh, escape =F, align="c") %>%
   kable_styling(bootstrap_options = c("striped", "hover"), full_width=F, position = "right", font_size=12)
```

## Fitting the model
```{r minerparms, invisible=T}
  miner_K <- 15
  miner_itrs <- 500
  miner_burn <- 180
  miner_alpha <- 0.1
  miner_beta_<- colSums(dtm) / sum(dtm) * 100
```

We can now fit the model. We choose `r miner_K` topics, again with a relatively diffuse prior for topics over documents ($alpha =$ `r miner_alpha`)[^1]  and an asymmetric prior for word over topics.

[^1]: a high alpha value means that each document is likely to cotain a mixture of most of the topics and not any single topic specifically. a low alpha value means that is it more likely that a document may contain mixture of just a few, or even only one, of the topics.

```{r minermodfit, eval=F}
  model <- FitLdaModel(dtm = dtm,
                       k = miner_K,
                       iterations = miner_itrs,
                       burnin = miner_burn,
                       alpha = miner_alpha,
                       beta  = miner_beta_,
                       optimize_alpha = TRUE,
                       calc_likelihood = FALSE,
                       calc_coherence = TRUE,
                       calc_r2 = FALSE)
```

<center>
### The model is not calculated within this notebook, see here. ###
</center>

```{r minermodload}
  load("./inputs/good_model.Rdata")
  model <- good_mod
  rm(good_mod)
```

The output from the model is an S3 object of class lda\_topic\_model.
It contains several objects:

* The most important are three matrices:
  * theta gives $P(\text{topic}_{k}\mid\text{document}_{d})$;
  * phi gives   $P(\text{word}_{v}\mid\text{topic}_{k})$;
  * gamma gives $P(\text{topic}_{k}\mid\text{word}_{v})$;

* data is the DTM or TCM used to train the model;
* alpha and beta are the Dirichlet priors for topics over documents and words over topics, respectively;
* log\_likelihood is $P(\text{words}\mid\text{topics})$ at each iteration;
* coherence gives the probabilistic coherence of each topic;
* r2 is the R-squared of the model given the data.

Let's explore the topic's properties
```{r topexp}
  model$top_terms <- GetTopTerms(phi = model$phi, M = 5)
  model$prevalence <- colSums(model$theta) / sum(model$theta) * 100
  model$labels <- LabelTopics(assignments = model$theta > 0.05,
                              dtm = dtm,
                              M = 1)
  model$summary <- data.frame(topic = rownames(model$phi),
                              label = model$labels,
                              coherence = round(model$coherence, 3),
                              prevalence = round(model$prevalence,3),
                              top_terms = apply(model$top_terms, 2, function(x){
                                paste(x, collapse = ", ")
                              }),
                              stringsAsFactors = FALSE)

  model$summary %>% mutate(coherence=  cell_spec(coherence, color = "white",
                           bold = T,
                           background = spec_color(order(coherence), end=0.9, option="B", direction=-1)
                           )) %>%
  kable(escape =F, align="c") %>%
   kable_styling(bootstrap_options = c("striped", "hover"), full_width=F, position = "center")
```

* The coherence terms measures how associated words are in a topic, controlling for statistical independance.
* The prevalence measures the frequency of each topic in the entire corpus

## Topics and papers interactions
Now that we have probability distribution for words in each topic, we can calculate similarity among topics using the [Hellinger distance](https://en.wikipedia.org/wiki/Hellinger_Distance).
The distance has been used as the base for [ward.D](https://en.wikipedia.org/wiki/Ward%27s_method) agglomerative hiearchical clustering (dendrogram on the left).
Expert knowledge has been used to decide to cut the tree in three groups: these re the three ``fields'' on which we will discuss our results.

```{r tophelldist}
  top.dist.mat <- CalcHellingerDist(model$phi) %>% as.dist
  nam <- model$summary$label_1
  names(top.dist.mat) <- paste(rownames(model$summary), " - ", nam, ifelse(duplicated(nam),"_bis","") , sep="")
```

```{r tophelltree}
  top.dendri <- hclust(top.dist.mat, "ward.D")
  top.clustering <- cutree(top.dendri, 3)
  a<- top.dendri %>% as.dendrogram %>% 
       set("branches_k_color", value = c( "#FF6600FF","#440154FF", "#21908CFF"), k = 3) %>%
       set("labels_cex", value=0.75) %>%
       hang.dendrogram %>%
       as.ggdend %>%
       ggplot
```

```{r paphelldist}
  pap.dist.mat <- CalcHellingerDist(model$theta) %>% as.dist
  pap.dr <- pap.dist.mat %>%
   stats::cmdscale(., k=2) %>%
   as_tibble() %>%
   rename(x=V1,y=V2)
```

```{r paphellmds}
  innet2 <- apply(model$theta, 1, function(x) {
                 names(x)[order(x, decreasing = TRUE)][1]
       })
  innet3 <- top.clustering[innet2]
  b <- ggplot(pap.dr %>% mutate(topw=innet3)) +
    aes(x=x, y=y, col=factor(topw)) +
    geom_point(size=2.5, alpha=0.75) +
  scale_color_manual(name="field",values = c("#440154FF", "#21908CFF", "#FF6600FF")) +
  theme_void() +
  scale_x_reverse()
```

```{r gridplot, warning=FALSE}
grid.arrange(a,b, ncol=2)
```
We use Hellinger distance also for calculating the distance between document pairs, based on the probability distribution for each topic in each paper.
Now We can performed a metric dimensional scaling on the papers' distance matrix, and display its first two dimensions as scatter plot (plot on the right).

## Alluvial plot
Finally, we display the connection between words and fields through topics using an alluvial diagram.
```{r papnet2, warning=F, message=F}
  blippo <- table(innet2) %>%
    enframe(name = "topic") %>%
    rename(count = "value") %>%
    left_join(top.clustering %>%
              enframe(name = "topic")) %>%
    rename(Fields= "value") %>%
    mutate(count = as.vector(count),
           name = c("IS", "ET", "EP", "ES", "MPA", "FI",
                    "SD", "MP", "BGC", "MR", "HAB", "CR",
                    "AQ", "CC", "FS"),
           Fields = factor(Fields,
                          labels = c("Conservation", "Management", "Impacts"))
           )
  blippo$exp_lab <- c("Invasive Species","Eco-tourism","Eco-physiology","Ecosystem Services","Marine Protected Areas","Fisheries","Sustainable Development","Marine Policy","Bigeochemical Cycles","Marine Resources","Harmul Algal Blooms","Coral Reefs","Aquaculture","Climate Change","Fish species")
  blupi <- blippo %>%
    left_join(model$summary) %>%
    select(-coherence,-prevalence) %>%
    separate_rows(top_terms, sep=",") %>%
    select(Fields,topic,exp_lab,top_terms, count) %>%
    mutate(top_terms=trimws(top_terms)) %>%
    mutate(topic=as.integer(gsub("t_","",topic))) %>%
    arrange(topic) %>%
    unite(topic,topic,exp_lab,sep=" - ") %>%
    mutate(topic=factor(topic, levels=unique(topic)))
  
  icol <- blupi$Fields %>% as.factor %>% as.integer
  tavol <- c(Impacts="#440154FF", Management="#21908CFF", Conservation="#FF6600FF")
  alluvial(blupi[,-4], freq=blupi$count, alpha=0.5, col=tavol[icol], cex=0.5, border = NA, blocks=FALSE)
```

<center>
## The next notebook will analyse Scopus data
</center>

```{r}
```


